---
output: 
    pdf_document
always_allow_html: yes
params:
  start: !r Sys.Date()-1
  stop: !r Sys.Date()-1
---

```{r setup, echo=FALSE}

# libraries
suppressPackageStartupMessages(library(knitr))
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(leaflet))
suppressPackageStartupMessages(library(rgdal))
suppressPackageStartupMessages(library(htmltools))
suppressPackageStartupMessages(library(htmlwidgets))
suppressPackageStartupMessages(library(maptools))
suppressPackageStartupMessages(library(lubridate))
suppressPackageStartupMessages(library(oce))
suppressPackageStartupMessages(library(leaflet.extras))

# designate prefix for paths
pre = '../'

# for testing
# params=data.frame(start = as.Date('2019-04-04'),
                  # stop = as.Date('2019-04-03'))

source(paste0(pre,'R/functions.R'))

# read in data
obs = readRDS(paste0(pre, 'data/processed/observations.rds'))
trk = readRDS(paste0(pre, 'data/processed/tracks.rds'))
load(paste0(pre, 'data/processed/tss.rda'))
load(paste0(pre, 'data/processed/gis.rda'))
bb_file = paste0(pre, 'data/raw/gis/canadian_boundary/canadian_boundary.csv')

# define species list
spp_list = c('right', 'fin', 'sei', 'humpback', 'blue')

# rename time limits
t0 = params$start
t1 = params$stop

# subset observations
obs = obs %>%
  filter(date >= t0 & date <= t1 & 
           species %in% spp_list &
           score %in% c('definite visual','definite acoustic') &
           name != 'jasco_test' &
           !is.na(lat) & !is.na(lon))
obs = subset_canadian(obs, bb_file = bb_file)
  
# subset tracks
trk = trk %>%
  filter(date >= t0 & date <= t1 & 
           name != 'jasco_test' &
           !is.na(lat) & !is.na(lon))
trk = subset_canadian(trk, bb_file = bb_file)

# create data to/from message
if(t0 == t1){
  data_range = params$stop
} else {
  data_range = paste0(params$start, ' to ', params$stop)
}
```

\begin{center}
	\huge \textbf{WhaleMap Summary Report} \\
	\huge \textbf{`r data_range`} \\
	\bigskip
	\large 
	\textbf{Report generated on:} `r Sys.time()`\\
	\textbf{Produced by Dalhousie MEOPAR-WHaLE project and DFO Science - Maritimes Region} \\
\end{center}

\hrulefill

#### Figure 1: Locations of large whale observations

```{r map, echo=FALSE, warning=FALSE, fig.height=8, fig.width=8}

# setup -------------------------------------------------------------------

# define score color palette
pal = colorFactor(levels = spp_list, palette = c('red', 'blue', 'lightgray', 'green', 'darkslategrey'))

# define function to determine trackline color
getColor <- function(trk) {
  if(trk$platform[1] == 'slocum') {
    "blue"
  } else if(trk$platform[1] == 'plane') {
    "#8B6914"
  } else if(trk$platform[1] == 'vessel'){
    "black"
  } else {
    "darkgrey"
  }
}

# combine all lat/lon to help determine plot limits
inLat = c(obs$lat, trk$lat)
inLon = c(obs$lon, trk$lon)

# basemap -----------------------------------------------------------------

# start basemap
map <- leaflet(options = leafletOptions(zoomControl = FALSE, attributionControl = FALSE)) %>%
  
  # add ocean basemap
  addProviderTiles(providers$Esri.OceanBasemap) %>%
  
  # add legend
  addLegend(position = "topleft", 
            title = 'Species',
            opacity = 0.5,
            pal = pal,
            values = spp_list) %>%
  
  # add extra map features
  addScaleBar(position = 'topright') %>%
  
  # use NOAA graticules
  addWMSTiles(
    "https://gis.ngdc.noaa.gov/arcgis/services/graticule/MapServer/WMSServer/",
    layers = c("1-degree grid", "5-degree grid"),
    options = WMSTileOptions(format = "image/png8", transparent = TRUE),
    attribution = NULL)

# center on focal region
if(length(inLon)!=0){
  map <- fitBounds(map = map, 
                   lng1 = min(inLon, na.rm = T), 
                   lng2 = max(inLon, na.rm = T), 
                   lat1 = min(inLat, na.rm = T), 
                   lat2 = max(inLat, na.rm = T), 
                   options = list(maxZoom=7)) 
}

# plot polygons -----------------------------------------------------------

# add mpas
map <- map %>%
  addPolygons(data=critical_habitat_zone, fill = T, fillOpacity = 0.25, stroke = T, smoothFactor = 0,
              dashArray = c(5,5), options = pathOptions(clickable = F),
              weight = 1, color = 'darkgreen', fillColor = 'darkgreen')

# plot shipping lanes
map <- map %>%
  addPolylines(tss_lines$lon, tss_lines$lat,
               weight = .5,
               color = 'grey',
               smoothFactor = 0,
               options = pathOptions(clickable = F)) %>%
  addPolygons(tss_polygons$lon, tss_polygons$lat,
              weight = .5,
              color = 'grey',
              fillColor = 'grey',
              smoothFactor = 0,
              options = pathOptions(clickable = F))

# add tracks --------------------------------------------------------------

# set up track plotting
trk.df <- split(trk, trk$id)

# add lines
names(trk.df) %>%
  purrr::walk( function(df) {
    map <<- map %>%
      addPolylines(data=trk.df[[df]],opacity = 0.3,
                   lng=~lon, lat=~lat, weight = 2,
                   smoothFactor = 0, color = getColor(trk.df[[df]]))
  })

# add definite detections/sightings ---------------------------------------

map <- map %>% addCircleMarkers(data = obs, ~lon, ~lat,
                                radius = 3, fillOpacity = 0.5, stroke = T, col = 'black', weight = 0.5,
                                fillColor = pal(obs$species))

# show map
if(length(inLon)!=0){
  map
} 
```

```{r map_text, echo=FALSE, results='asis'}

# show map text
if(length(inLon)!=0){
  cat('For interactive map, go to: https://whalemap.ocean.dal.ca/WhaleMap/ \n')
  cat('\\clearpage \n')
} else {
  cat('*No effort or observations were recorded during this period*')
}

```

#### Table 1: Number of large whale observations

```{r spp_table, echo=FALSE, results='asis'}

# count numbers of all species sighted
vis = obs[obs$score=='definite visual',]
vis = obs[!is.na(obs$number),]

if(nrow(vis)!=0){
  vis = aggregate(number ~ species, data = vis, FUN = sum)
} else {
  vis = data.frame(species = levels(obs$species), number = rep(0, length(levels(obs$species))))
}

# count acoustic detections
aco = as.data.frame(table(obs$species[obs$score=='definite acoustic']))

# merge sightings and acoustics
spp_table = merge(vis, aco, by.x = 'species', by.y = 'Var1', all.y = TRUE)

# rename columns
colnames(spp_table) = c('species', 'visual_counts', 'acoustic_detections')

# replace zeros
spp_table$visual_counts[is.na(spp_table$visual_counts)] = 0

# sort table alphabetically
spp_table=spp_table[order(as.character(spp_table$species), decreasing = FALSE),]

# display table
kable(spp_table, row.names = FALSE, align = 'c')  

```

*Visual counts only include sightings where numbers of whales were recorded, and may include duplicates*

#### Table 2. Platforms with recorded survey effort

```{r effort, echo=FALSE, results='asis'}

# display table or message
if(nrow(trk)==0){
  cat('*No platforms recorded effort during this period*')
} else {
  platform_table = aggregate(platform ~ name, data = trk, FUN = unique)
  kable(platform_table, row.names = FALSE, align = 'c')
}

```

#### Table 3: North Atlantic right whale sightings

```{r narw_sig, echo=FALSE, results='asis'}

# select species
sig = obs[obs$species == 'right' & obs$score == 'definite visual',]

# format table
sig = sig[,c('date','time', 'lat', 'lon', 'number', 'platform', 'name')]

# sort by time
sig = sig[order(sig$time),]

# format timestamps
sig$time = format(sig$time, '%H:%M:%S')

# display table or message
if(nrow(sig)==0){
  cat('*No right whales were sighted*')
} else {
  kable(sig, row.names = FALSE, align = 'c')  
}

```

#### Table 4: North Atlantic right whale acoustic detections

```{r narw_det, echo=FALSE, results='asis'}

# select species
det = obs[obs$species == 'right' & obs$score == 'definite acoustic',]

# format table
det = det[,c('date','time', 'lat', 'lon', 'platform', 'name')]

# sort by time
det = det[order(det$time),]

# format timestamps
det$time = format(det$time, '%H:%M:%S')

# display table or message
if(nrow(det)==0){
  cat('*No right whales were acoustically detected*')
} else {
  kable(det, row.names = FALSE, align = 'c')  
}

```
